---
title: "Recommendation Engine"
author: "Pramod Verma"
date: "7 December 2017"
output: html_document
---


**Load the required package into the memory**
```{r, warning=FALSE, message=FALSE}
library("recommenderlab")
```


**Creating small dataset**
```{r}
m <- matrix(sample(c(as.numeric(0:5), NA), 50,
                   replace=TRUE, prob=c(rep(.4/6,6),.6)), ncol=10,
            dimnames=list(user=paste("u", 1:5, sep=""),
                          item=paste("i", 1:10, sep="")))
m
```


**With coercion, the matrix can be easily converted into a realRatingMatrix object which stores
the data in sparse format (only non-NA values are stored explicitly; NA values are represented by a dot)**
```{r}
r <- as(m, "realRatingMatrix")
r
```

**An important operation for rating matrices is to normalize the entries to, e.g., centering to
remove rating bias by subtracting the row mean from all ratings in the row. This is can be
easily done using normalize()**
```{r}
r_m <- normalize(r)
r_m
image(r, main = "Raw Ratings")
image(r_m, main = "Normalized Ratings")
```

**We will use the data set Jester5k for the rest of this section. This data set comes with recommenderlab
and contains a sample of 5000 users from the anonymous ratings data from the Jester joke Recommender**
```{r}
data(Jester5k)
Jester5k

set.seed(1234)
r <- sample(Jester5k, 1000)
r
hist(getRatings(r), breaks=100, col = "green")
hist(getRatings(normalize(r)), breaks=100, col = "green")
```

**Next, we create a recommender which generates recommendations solely on the popularity of items**

```{r}
r <- Recommender(Jester5k[1:1000], method = "POPULAR")
```

**Here we create top-5 recommendation lists for two users who were not used to learn the model**
```{r}
recom <- predict(r, Jester5k[1001:1002], n=5)
recom
```


**We can get the best 3 recommendations for each list using bestN()**
```{r}
recom3 <- bestN(recom, n = 3)
recom3
```


**Many recommender algorithms can also predict ratings. This is also implemented using
predict() with the parameter type set to "ratings"**
```{r}
recom <- predict(r, Jester5k[1001:1002], type="ratings")
```

**Alternatively, we can also request the complete rating matrix which includes the original
ratings by the user**
```{r}
recom <- predict(r, Jester5k[1001:1002], type="ratingMatrix")
recom
```

#Evaluation of predicted ratings
```{r}
e <- evaluationScheme(Jester5k[1:1000], method="split", train=0.9, given=15, goodRating=5)
e
```

**We create two recommenders (user-based and item-based collaborative filtering) using the
training data.**
```{r}
r1 <- Recommender(getData(e, "train"), "UBCF")
r1
```


**Recommender of type 'UBCF' for 'realRatingMatrix' learned using 900 users**
```{r}
r2 <- Recommender(getData(e, "train"), "IBCF")
r2
```


**Next, we compute predicted ratings for the known part of the test data (15 items for each user) 
using the two algorithms**
```{r}
p1 <- predict(r1, getData(e, "known"), type="ratings")
p1
p2 <- predict(r2, getData(e, "known"), type="ratings")
p2
```


**Finally, we can calculate the error between the prediction and the unknown part of the test data**
```{r}
error <- rbind(
  UBCF = calcPredictionAccuracy(p1, getData(e, "unknown")),
  IBCF = calcPredictionAccuracy(p2, getData(e, "unknown")))

error
```


**Evaluation of a top-N recommender algorithm**
```{r}
scheme <- evaluationScheme(Jester5k[1:1000], method="cross", k=4, given=3, goodRating=5)
scheme
```


**Next we use the created evaluation scheme to evaluate the recommender method popular.
We evaluate top-1, top-3, top-5, top-10, top-15 and top-20 recommendation lists**
```{r}
results <- evaluate(scheme, method="POPULAR", type = "topNList", n=c(1,3,5,10,15,20))
results
getConfusionMatrix(results)[[1]]
avg(results)
plot(results, annotate=TRUE)
```


**plot the precision and recall**
```{r}
plot(results, "prec/rec", annotate=TRUE)
```

**Comparing recommender algorithms**
```{r}
set.seed(2016)
scheme <- evaluationScheme(Jester5k[1:1000], method="split", train = .9,
                           k=1, given=-5, goodRating=5)
scheme

algorithms <- list(
  "random items" = list(name="RANDOM", param=NULL),
  "popular items" = list(name="POPULAR", param=NULL),
  "user-based CF" = list(name="UBCF", param=list(nn=50)),
  "item-based CF" = list(name="IBCF", param=list(k=50)),
  "SVD approximation" = list(name="SVD", param=list(k = 50)))
```

**run algorithms**
```{r}
results <- evaluate(scheme, algorithms, type = "topNList", n=c(1, 3, 5, 10, 15, 20))
plot(results, annotate=c(1,3), legend="bottomright")
plot(results, "prec/rec", annotate=3, legend="topleft")

```

#End of File
Reference: https://cran.r-project.org/web/packages/recommenderlab/vignettes/recommenderlab.pdf

